---
title: "Homework 1: Linear Regression"
author: "Chamroeun Chhay"
output: html_document
editor_options: 
  chunk_output_type: console
---

There are six questions (30 total points) in this assignment. The minimum increment is 1 point. Please type in your answers directly in the R Markdown file. After completion, **successfully** knitr it as an html file. Submit [**both**]{style="color:red"} the html file and the R Markdown file via Canvas. Please name the R Markdown file in the following format: LastName_FirstName_HW1.Rmd, e.g. Zhao_Zifeng_HW1.Rmd.

## Used Car Dataset [9 points]

The used car dataset is the one we analyzed in class. Let's read in the data stored in `UsedCar.csv`.

```{r chunk1}
usedcar <- read.csv('~/Downloads/UsedCar.csv')

```

### **Q1 [2 points]**

Since the dependent variable `Price` is severely right-skewed, create a log-transformation of `Price` and store it as a new variable named `Log_price` within the data.frame `total_data`.

```{r Q1}
usedcar$log_price <- log(usedcar$Price + 1)

```

### **Q2 [7 points]**

Fit a linear regression model of Log_price w.r.t. two predictors Age and Mileage, name it `lm_short`.

```{r Q2}
lm_short <- lm(log_price ~ Age + Mileage, data = usedcar)

```

**Q2(a) [2 points]** What is the $R^2$ of `lm_short`? What are the (in-sample) MAE and RMSE of `lm_short` at the *original scale*?

```{r Q2(a)}
summary(lm_short)

library(forecast)

original_scale <- exp(lm_short$fitted.values) - 1 

accuracy(original_scale, usedcar$Price)

```

Answer: The $R^2$ of `lm_short` is .8055. The in-sample MAE and RMSE of `lm_short` at the *original scale* is MAE (1065.2) and RMSE (1526.21).

**Q2(b) [2 point]** What is the estimated coefficient of `lm_short` for Age and Mileage?

```{r Q2(b)}
lm_short$coefficients
```

**Q2(c) [3 points]** How should we interpret the estimated coefficient of Age at the log-scale and the original scale of Price?

Answer: the log-scale shows that the Log Price decrease by 1.217e-02 for each unit increase in Age. For the original scale, the original price would decrease by (1.216931e-02 \* 100) % or 1.22% decrease for each unit increase in Age.

## Car Seat Sales Dataset [21 points]

The car seat sales dataset contains sales of child car seats at 400 different stores and the data is stored in `Carseats.csv`. It contains 9 variables, `Sales`, `CompPrice`, `Income`, `Advertising`, `Population`, `Price`, `ShelveLoc`, `Age` and `Urban`. We would like to build a linear regression model to predict `Sales` at a planned new store. The data description is as follows.

-   `Sales`: Unit sales (in thousands) at each location
-   `CompPrice`: Price charged by competitor at each location
-   `Income`: Community income level (in thousands of dollars)
-   `Advertising`: Local advertising budget for company at each location (in thousands of dollars)
-   `Population`: Population size in region (in thousands)
-   `Price`: Price company charges for car seats at each site
-   `ShelveLoc`: A factor with levels Bad, Good and Medium indicating the quality of the shelving location for the car seats at each site
-   `Age`: Average age of the local population
-   `Urban`: A factor with levels No and Yes to indicate whether the store is in an urban or rural location

### **Q4 [2 points]**

Which variable is the dependent variable? Which predictors are categorical variables?

Answer: The dependent variable is the `Sales`. The categorical variables are `ShelveLoc` and `Urban`. 

### **Q5 [9 points]**

Let's read in the data and perform visualization to get a better sense of the data.

**Q5(a) [2 points]** Correctly read in the data stored at `Carseats.csv`.

```{r Q5(a)}
carseats <- read.csv('~/Downloads/Carseats.csv')

```

**Q5(b) [2 points]** Produce a scatterplot between `Sales` and `Price`. What is the general pattern from the scatterplot?

```{r Q5(b)}
plot(carseats$Price, carseats$Sales, xlab = 'Price', ylab = 'Sales')

```

Answer: There seems to be a negative relationship between `Price` and `Sales`. The scatter plot shows that as `Price` increases, `Sales` decreases.

**Q5(c) [2 points]** Produce a boxplot between `Sales` and `ShelveLoc`. What is the general pattern from the boxplot?

```{r Q5(c)}
boxplot(carseats$Sales ~ carseats$ShelveLoc,xlab = 'ShelveLoc', ylab = 'Sales', main = 'Sales vs ShelveLoc')

```

Answer: The boxplot shows that the better the `ShelveLoc`, the higher number of `Sales` there are. 

**Q5(d) [3 points]** Produce a histogram of `Sales`. Is `Sales` severely right-skewed?

```{r Q5(d)}
hist(carseats$Sales)

```

Answer: The histogram does not show that the data is skewed in any direction. There seems to be a normal distribution in the data. 

### **Q6 [10 points]**

**Q6(a) [4 points]** Fit a linear regression model of the original scale `Sales` w.r.t. all the predictors available in the dataset, name it `lm_full`. Take a look at the summary of `lm_full`.

```{r Q6(a)}
lm_full <- lm(Sales ~., data = carseats)
summary(lm_full)

```

**Q6(b) [2 points]** What is the estimated coefficient of `lm_full` for CompPrice and Price? How should we interpret the estimated coefficients?

```{r Q6(b)}
lm_full$coefficients

```

Answer: The estimated coefficient for `CompPrice` is 0.09268 and for `Price` is  -0.0954. This means that the for each unit increase in `CompPrice`, the `Sales` number would increase by 0.0926. For the `Price`, if the `Price` increase by a unit, the number of `Sales` decreases by 0.0954. 

**Q6(c) [2 points]** Which predictor(s) are not statistically significant in the model?

Answer: The predictors that are not statistically significant in the model are `Population` and `UrbanYes`, since their p-values are greater than .05. 

**Q6(d) [2 points]** What is the $R^2$ of `lm_full`? What are the (in-sample) MAE and RMSE of `lm_full`?

```{r Q6(d)}
summary(lm_full)

accuracy(lm_full$fitted.values, carseats$Sales)
```

Answer: The $R^2$ of `lm_full` is  0.8726. The RMSE is 1.006678 and the MAE is 0.805898.

